### 2025-04-29 lecturenotes_aima203/ch2_aima203.tex

\documentclass[aima203_lecturenotes_ku.tex]{subfiles}

\setcounter{chapter}{1}
\begin{document}

\chapter{Root Finding}
\section{Introduction}
Roots of an equation
\begin{equation}
  \label{eq:8}
f(x) = 0
\end{equation}
are the zeros, of \(f\), which means, the values of \(x\) that makes the value of \(f\) zero. Basically equations are categorized into two. If \(f\) is a polynomial then the Equation-\ref{eq:8} is a \textbf{polynomial equation} and if \(f\) is a non-polynomial then Equation-\ref{eq:8} is a \textbf{transcendental equation}. For the polynomial equations following results hold:
\begin{enumerate}
\item Every polynomial equation of degree \(n\) has at most \(n\) real roots.
\item If \(n\) is odd then, the polynomial equation has at least one real root whose sign is opposite to that of the last term.
\item If \(n\) is even and the constant term is negative, then the equation has at-least one positive root and at-least one negative root.
\item The imaginary roots occurs in a pair (conjugate-pair). If the coefficients of \(f\) are rationals then, the irrational roots occurs in pairs (conjugate-pair).

\item \textbf{Descartes' Rule of Signs}
  \begin{enumerate}
  \item[a).] A polynomial equation cannot have more number of positive real roots than the number of changes of signs in the coefficients of \(f(x)\).
  \item[b).] A polynomial equation cannot have more number of negative real roots than the number of changes of signs in the coefficients of \(f(-x)\).
  \end{enumerate}
\end{enumerate}

\section{Bisection Method}
\begin{mdframed}[style=myframe]
  \begin{theorem}[Bolzano's Theorem]
    \label{Bolzano}
  If f(x) is continuous in \([a,b]\), and if \(f(a)\) and \(f(b)\) are of opposite signs, then \(f(c)=0\) for at least one number \(c \in (a,b)\).
\end{theorem}
\end{mdframed}
The Bisection method is based on Theorem-\ref{Bolzano}. The word ``bisection'' means ``half''. Using this method the root \(c\) of \(f\) is given by \(\displaystyle c \approx \frac{a+b}{2}\). Let \(\displaystyle x_1=\frac{a+b}{2}\). If \(f(x_1) \neq 0\) then, the root, \(c\) lies either in \([a,x_1]\) or in \([x_1,b]\). If \(f(a)f(x_1) <0\) then, \(c\) lies in \([a,x_1]\) else, it lies in \([x_1,b]\). \\[4mm]
At each step of this method, the given interval is bisected, so the length of the interval is halfed. At \(nth\) step the length of the interval is \(\displaystyle \frac{|b-a|}{2^n}\). If the tolerance of the given approximation is \(\epsilon\) then we must have \(\displaystyle \frac{|b-a|}{2^n} \leq \epsilon\). And the number of steps required to reach this accuracy is \(\displaystyle n \geq {log_2(|b-a|)}\).

\subsection{Procedure}
\begin{enumerate}
\item Choose two real numbers \(a\) and \(b\) such that \(f(a)f(b) < 0\).
\item Set \(\displaystyle x_0 = 0\) and \(\displaystyle x_1= \frac{a+b}{2}\).
\item Do \\
  \(\displaystyle \epsilon_r = \left | \frac{x_0 -x_1}{x_0} \right |\) \\[1mm]
  If \(\epsilon_r < tolerance\) then \(root = x_1\), \\
  else\\
  \(x_0=x_1\) and if \(f(a)f(x_1) < 0\) then \(\displaystyle x_1= \frac{a+x_1}{2}\) \\
  if \(f(x_1)f(b) < 0\) then \(\displaystyle x_1= \frac{x_1+b}{2}\).
\end{enumerate}

\subsection{Exercise}
Using Bisection Method:
\begin{enumerate}
\item Find a root of \(f(x)=x^3-x-1=0\), correct to \(4\) decimal places
\end{enumerate}

\section{Iteration Method}
Steps:
\begin{enumerate}
\item Re-write the given equation \(f(x)=0\) in the form \(x= \phi (x)\).
  This equation is of \textbf{iterative-type}. Meaning we can substitute a value of \(x\) in \(\phi (x)\) to get another value of \(x\), and continue this process to get the desired value of \(x\) if the iteration is of convergent one.
\item Choose an initial root of \(f\), \(x_0\).
\item \(x_1=\phi (x_0)\), \(x_2= \phi(x_1)\) and so on.
\end{enumerate}
The sequence \(x_0, x_1, x_2, ...\) may not converge to a definite number. But if the sequence converges to a definite number \(\zeta\), then \(\zeta\) is a root of the given equation.

\subsection{Exercise}
Using Iteration Method:
\begin{enumerate}
\item Find a root of \(2x-3-cosx=0\), correct to \(3\) decimal places
\end{enumerate}

\section{Newton-Rapshon's Method}
Steps:
\begin{enumerate}
\item Choose an initial guess solution of the given equation \(f(x)=0\), \(x_0\).

\item Let \(x_1\) be a solution, which is more close to the exact solution of \(f(x)=0\). Then Using Taylor's expansion of \(f\) about \(x_0\): \[f(x_1)=f(x_0) + (x_1-x_0)f'(x_0) + (x_1+x_0)^2f''(x_0) + ... = 0\]
  Neglecting the second and higher order derivatives, we get
  \begin{equation}
    \label{taylor}
  f(x_0) + (x_1-x_0)f'(x_0)=0
  \end{equation}
  \begin{footnotesize}
    The equation-\ref{taylor} is a linear equation, so this is an linear approximation. This equation is infact the tangent to the curve of the function \(f(x)\) at $(x_0, f(x_0)$. And it is the point $x_1$ where the tangent meets the $x-axis$. So, the next approximation after $x_0$ by Newton-Rapshon's method is the point on $x-axis$, where the tangent to the $f$ at $x_0$ meets the $x-axis$. This point can be solved as follows:
   \end{footnotesize}
  \begin{align*}
    x_1-x_0 & = - \frac{f(x_0)}{f'(x_0)} \\
    x_1 &= x_0 - \frac{f(x_0)}{f'(x_0)}
  \end{align*}

\item Successive approximation are given by \(x_2, x_3, x_4, ...\), where
  \(\displaystyle x_{n+1} = x_n - \frac{f(x_n)}{f'(x_n)}\)
\end{enumerate}

\subsection{Exercise}
Using Newton-Rapshon's Method:
\begin{enumerate}
\item Find a root of \(f(x)=xe^x-1=0\), correct to \(4\) decimal places
\end{enumerate}

\section{Secant Method}
In Newton-Rapshon's method we use a tangent to the curve to get close to the root of the function. So, Newton-Rapshon's method requires the evaluation of derivatives of the function, which may not always exit. So we replace the tangent, with a secant to approximate the root of the function. \\[2mm]
Steps:
\begin{enumerate}
\item Choose two initial guess solutions of the given equation \(f(x)=0\), \(x_{-1}\) and \(x_0\).

\item The slope of the secant is \(\displaystyle \frac{f(x_0)-f(x_{-1})}{x_0 - x_{-1}}\).

\item  Then equation of the line passing through the points of given by the two initial guesses is \(\displaystyle f(x)-f(x_0)= \frac{f(x_0)-f(x_{-1})}{x_0 - x_{-1}} (x - x_0) \).

\item $x_1$ is the point where the secant meets the $x-axis$ so, $f(x_1)=0$. This gives,
  \begin{align*}
    0-f(x_0) &= \frac{f(x_0)-f(x_{-1})}{x_0 - x_{-1}} (x_1 - x_0) \\
    x_1 -x_0 &= - \frac{x_0 - x_{-1}}{f(x_0)-f(x_{-1})} f(x_0) \\
    x_1 &= x_0 - \frac{x_0 - x_{-1}}{f(x_0)-f(x_{-1})} f(x_0)
  \end{align*}

\item This generalizes to
  \begin{equation}
    \label{eq:secant}
    x_{n+1} = x_n - \frac{x_n - x_{n-1}}{f(x_n)-f(x_{n-1})} f(x_n)
  \end{equation}
\end{enumerate}
You can get this relation-\ref{eq:secant} just by plugging $f'(x_n)=\frac{f(x_n)-f(x_{n-1})}{x_n - x_{n-1}}$ as the slope of the tangent in Newton Rapshon's method is just approximated by the slope of the secant in Secant method.

\section{System of Non-linear equations}
For now we consider only a system of two equations. Let a system of two equations be
\begin{equation}
  \label{system}
  f(x,y)=0, \hspace{1cm} g(x,y)=0
\end{equation}
\subsection{Method of Iteration}
First we assume that the sytem of equations~\ref{system} may be written in the form
\begin{equation}
  \label{iterate}
  x=F(x,y), \hspace{1cm} y=G(x,y)
\end{equation}
where the function $F$ and $G$ satisty the following conditions in a \textbf{closed} neighborhood of $R$ of the root $(\alpha, \beta)$:
\begin{enumerate}
\item[i)] $F$ and $G$ and their firt partial derivatives are continuous in $R$, and
  \item[ii)] $\displaystyle \left | \frac{\partial F}{\partial x} \right | + \left | \frac{\partial F}{\partial y} \right | < 1$ and $\displaystyle \left | \frac{\partial G}{\partial x} \right | + \left | \frac{\partial G}{\partial y} \right | < 1$, for all $(x,y)$ in $R$.
\end{enumerate}
If $(x_0,y_0)$ is an initial approximation to the root $(\alpha, \beta)$, then Equations~\ref{iterate} give the sequence
\begin{equation}
\begin{gathered}
  x_1 = F(x_0,y_0), \hspace{2cm} y_1 = G(x_0,y_0) \\
  x_2 = F(x_1,y_1), \hspace{2cm} y_2 = G(x_1,y_1) \\
  ... \\
  x_{n+1} = F(x_n,y_n), \hspace{2cm} y_{n+1} = G(x_n,y_n)
\end{gathered}
\end{equation}
For faster convergence, recently computed values of $x_i$ may be used in the evaluation of $y_i$ in Equations. Above conditions are sufficient for convergence and in the limit we obtain,
\begin{equation}
  \alpha = F(\alpha, \beta) \hspace{5mm} and \hspace{1cm} \beta=G(\alpha, \beta)
\end{equation}
Hence $(\alpha, \beta)$ is the root of the system~\ref{system}.

\subsection{Newton-Raphson Method}
Let $(x_0,y_0$ be an initial approximation to the root of the system of equations in two variables~\ref{system}. If $(x_0+h, y_0+k$ is the root of the system, then we must have
\begin{equation*}
  f(x_0+h,y_0+k) = 0 \hspace{1cm} g(x_0+h,y_0+k) = 0
\end{equation*}
Assuming that $f$ and $g$ are sufficiently differentiable, we expand both of these functions by Taylor's series to obtain
\begin{equation*}
  \begin{gathered}[t]
    f_0 +h\frac{\partial f}{\partial x_0} + k\frac{\partial f}{\partial y_0}... = 0 \\
    g_0 +h\frac{\partial g}{\partial x_0} + k\frac{\partial g}{\partial y_0}... = 0
  \end{gathered}
\end{equation*}
where, \hspace{5mm} $\displaystyle \frac{\partial f}{\partial x_0} = \left [\frac{\partial f}{\partial x} \right ] _{x=x_0}$, $f_0 = f(x_0,y_0)$, etc \\[1mm]
Neglating the second and higher-order derivatives terms, we get,
\begin{equation}
  \label{taylor}
  \begin{gathered}[b]
    h\frac{\partial f}{\partial x_0} + k\frac{\partial f}{\partial y_0}... = -f_0 \\
     h\frac{\partial g}{\partial x_0} + k\frac{\partial g}{\partial y_0}... = g_0
  \end{gathered}
\end{equation}
The system of equations~\ref{taylor} possesses a unique solution if
$$ D =
\left \vert \begin{matrix}
  \frac{\partial f}{\partial x_0}  & \frac{\partial f}{\partial y_0} \\[2mm]
  \frac{\partial g}{\partial x_0}  & \frac{\partial g}{\partial y_0}
\end{matrix} \right \vert  \neq 0
$$
By Cramer's rule
\begin{equation}
  h= \frac{1}{D}\, \left \vert \begin{matrix}
  -f_0  & \frac{\partial f}{\partial y_0} \\[2mm]
  -g_0  & \frac{\partial g}{\partial y_0}
\end{matrix} \right \vert \hspace{5mm} and \hspace{5mm} k= \frac{1}{D}\, \left \vert \begin{matrix}
  \frac{\partial f}{\partial y_0} & -f_0  \\[2mm]
 \frac{\partial g}{\partial y_0} &  -g_0
\end{matrix} \right \vert
\end{equation}
The new approximations are, therefore
\begin{equation}
  x_1 = x_0 +h \hspace{5mm} and \hspace{5mm} y_1 = y_0 + k
\end{equation}

\subsection{Exercise}
\begin{enumerate}
\item Find a real root of the system: $y^2-5y+4=0$ and $3x^2y-10x+7 = 0$ correct to $4$ decimal places using initial approximation $(0,0)$.

\item Solve the system: $x^2+y = 11$, $x+y^2 = 7$.

\item Solve the system: $x^2-y^2 = 4$, $x^2+y^2=16$.
\end{enumerate}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: