### 2025-04-29 lecturenotes_aima104/title_ai104.tex

\vspace*{2cm}
{\begin{center}
  {\Huge
    {\bfseries {\thetitle}}} \\
        \vspace{5mm}
        {\large AIMA 104}
  \vspace{3cm}

\begin{figure}[h]
	\centering
	\includegraphics[height=4cm,width=4cm]{images/Kulogo.png}
\end{figure}

\vspace{1cm}

	{\large \bfseries Department of Artificial Intelligence, \\[1mm]
 Kathmandu University, Panchkhal, Kavre}

\vspace{2cm}
        \textit{Lecture notes}
	by\\
	{\bfseries  \theauthor}

\vspace{1cm}

\textbf{Text Book}:
\begin{enumerate}
\item David C. Lay, Linear Algebra and Its Applications (4th edition), Pearson.
\item Gilbart Strang, Introduction to Linear Algebra, Wellesley-Cambridge Press.
\end{enumerate}
    \vspace{1.5cm}

{\thedate}

\thispagestyle{empty}
\end{center}}

\clearpage