### 2025-05-30 question_papers/aima203_first_in_sem.tex
question paper

\documentclass[12pt]{exam}
\usepackage{preamble_question}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
 {\bfseries  {\Large Kathmandu University}} \\[1mm]
 First In-Semester Exam-2025\\[-1mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\vskip 5mm

\begin{minipage}{0.70\linewidth}
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 2mm
\rule{\textwidth}{1pt}
\hskip 15mm \textit{Write the answer according to the mark allocated to it}. \textit{For short questions, write short answers, for long questions write long answers}.

\vskip 10mm

\begin{questions}

\question Round off the digits 45.321 and 12.987 upto two decimal places and find the error in their product after rounding off. [2]

\question

\end{questions}
\end{document}