### 2025-05-30 question_papers/aima203_mcq_first_in_sem.tex

\documentclass[12pt]{exam}
\usepackage{preamble_question}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
  {\bfseries  {\large Kathmandu University}} \\
  First In-semester Examination-2025 \\[-2mm]
    \textbf{\small Department of Artificial Intelligence, Panchkhal}
\end{center}
\vspace{2mm}

\begin{minipage}{0.70\linewidth}
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 15 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 10
  \end{flushleft}
\end{minipage}
\vspace{-8mm}
\begin{center}
  \rule{\textwidth}{1pt}
  Roll No: \hspace{5.5cm} Marks Scored:
  \vskip -3mm
\rule{\textwidth}{1pt}
\end{center}

\begin{center}
  SECTION''A'' \hspace{5mm} [10 Q \(\times\) 1 = 10 marks] (\textit{Do any ten}).
\end{center}
Fill in the blanks by writing the most appropriate word(s) or symbol(s).
\begin{questions}
  \question Give an example of an approximate value: \rule{8cm}{0.15mm}.

  \question Write the number of significant digits in 20300.00  \rule{5cm}{0.15mm}.

  \question Write the precision of 45.203 \rule{5cm}{0.15mm}.

  \question Write the accuracy of 7953.2300 \rule{5cm}{0.15mm}.

\end{questions}

\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: