### 2025-06-08 question_papers/aima104_first_in_sem.tex
question aima104

\documentclass[12pt]{exam}
\usepackage{preamble_question}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
 {\bfseries  {\Large Kathmandu University}} \\[1mm]
 First In-Semester Exam-2025\\[-1mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\vskip 5mm

\begin{minipage}{0.70\linewidth}
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: I \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 104 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 2mm
\rule{\textwidth}{1pt}
\hskip 15mm \textit{Write the answer according to the mark allocated to it}. \textit{For short questions, write short answers, for long questions write long answers}.

\vskip 10mm

\begin{questions}
\question For a matrix $A$ of order $m \times n$. Show that $Null(A)$ is a subspace of $\mathbb{R}^n$. [2]

  \question Show that the transformation $T(x)=x+3$ is not linear. [2]

  \question Describe the transformation of the following matrix geometrically:
  $\begin{bmatrix}
    1 & 0\\
    0 & 0
  \end{bmatrix}$. [2]

\question Prove: The transformation $T: V \to W$ is linear, then $T(0)=0$. [2]

  \question Verify Rank Nullity Theorem: $\displaystyle \begin{bmatrix}
    1 & 2 & 3 \\
    6 & 5 &4 \\
    7 & 8 & 9
\end{bmatrix}.$ [4]

\question Show that the transformation that rotates each point in $\mathbb{R}^2$, about the origin, through an angle $\phi$ counterclockwise, is a linear transformation by find the standard matrix of the transformation. [4]

\question Solve: \hspace{2mm} $x_2+4x_3 = -5, \hspace{5mm} x_1+3x_2+5x_3 = -2, \hspace{5mm} 3x_1 + 7x_2 + 7x_3 = 6$. [4]

\end{questions}
\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: