### 2025-06-09 question_papers/aima203_first_in_sem.tex
interpolation

\documentclass[12pt]{exam}
\usepackage{preamble_question_small}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
 {\bfseries  { Kathmandu University}} \\[-1mm]
 First In-Semester Exam-2025\\[-2mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\begin{minipage}{0.70\linewidth}
  \small
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 1mm
\rule{\textwidth}{1pt}
\vskip -5mm
\begin{questions}

\question Round off the digits 45.321 and 12.987 upto two decimal places and find the error in their product after rounding off. [2]

\question Find a root of $x^3-4x-9=0$, using bisection method, with absolute error less than $10^2$. [4]

\question Describe error due to truncation. Find a root of \(f(x)=xe^x-1=0\), correct to \(2\) decimal places. [1+3]
\end{questions}

\vspace{3cm}
\begin{center}
 {\bfseries  { Kathmandu University}} \\[-1mm]
 First In-Semester Exam-2025\\[-2mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\begin{minipage}{0.70\linewidth}
  \small
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 1mm
\rule{\textwidth}{1pt}

\begin{questions}

\question Round off the digits 45.321 and 12.987 upto two decimal places and find the error in their product after rounding off. [2]

\question Find a root of $x^3-4x-9=0$, using bisection method, with absolute error less than $10^2$. [4]

\question Describe error due to truncation. Find a root of \(f(x)=xe^x-1=0\), correct to \(2\) decimal places. [1+3]

  \clearpage

\vspace{4cm}
\question Find a real root of the equation $x^2+y =1, x+y^2=7$, using the method of iteration, correct up-to 2 decimal places. [4]

\question Solve using LU-factorization method. $3x_1 - x_2 = -1, \;\;\; -x_1+3x_2-x_3 = 7, \;\;\; -x_2+3x_3 =7$. [4]
\vskip -5mm
\question Show that the matrix $ \small A=   \begin{bmatrix}
    25  & 24 & 10 \\
    66 & 78 & 37 \\
    92 & -73 & -80
  \end{bmatrix}
  $ is ill-conditioned.
\vspace{12cm}

\question Find a real root of the equation $x^2+y =1, x+y^2=7$, using the method of iteration, correct up-to 2 decimal places. [4]

\question Solve using LU-factorization method. $3x_1 - x_2 = -1, \;\;\; -x_1+3x_2-x_3 = 7, \;\;\; -x_2+3x_3 =7$. [4]
\vskip -5mm
\question Show that the matrix $ \small A=   \begin{bmatrix}
    25  & 24 & 10 \\
    66 & 78 & 37 \\
    92 & -73 & -80
  \end{bmatrix}
  $ is ill-conditioned.
\end{questions}
\end{document}