### 2025-06-11 question_papers/aima203_first_in_sem.tex
aima203_question

\documentclass[12pt]{exam}
\usepackage{preamble_question_small}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
 {\bfseries  { Kathmandu University}} \\[-1mm]
 First In-Semester Exam-2025\\[-2mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\begin{minipage}{0.70\linewidth}
  \small
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 1mm
\rule{\textwidth}{1pt}
\vskip -5mm
\begin{questions}

\question Round off the digits 45.321 and 12.987 upto two decimal places and find the error in their product after rounding off. [2]

\question Find a root of $x^3-4x-9=0$, using bisection method, with absolute error less than $10^2$. [4]

\question Describe error due to truncation. Find a root of \(f(x)=xe^x-1=0\), correct to \(2\) decimal places. [1+3]
\end{questions}

\vspace{3cm}
\begin{center}
 {\bfseries  { Kathmandu University}} \\[-1mm]
 First In-Semester Exam-2025\\[-2mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\begin{minipage}{0.70\linewidth}
  \small
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 1mm
\rule{\textwidth}{1pt}

\begin{questions}

\question Round off the digits 45.321 and 12.987 upto two decimal places and find the error in their product after rounding off. [2]

\question Find a root of $x^3-4x-9=0$, using bisection method, with absolute error less than $10^2$. [4]

\question Describe error due to truncation. Find a root of \(f(x)=xe^x-1=0\), correct to \(2\) decimal places. [1+3]
\end{questions}

  \clearpage

\vspace{4cm}
\begin{questions}
\question Solve using LU-factorization method. $3x_1 - x_2 = -1, \;\;\; -x_1+3x_2-x_3 = 7, \;\;\; -x_2+3x_3 =7$. [4]

\question Determine whether the matrix is ill-conditioned or not.$\small A=   \begin{bmatrix}
    25  & 24 \\
    66 & 78 \\
  \end{bmatrix}$. [2]

\question Using Lagrange's method estimate the value of $y(202)$ from the following data: \\[2mm]
  $(200, 5.2983), (203,5.3132), (206,5.3278), (210,5.3471)$. [4]
  \vspace{12cm}

 \question Solve using LU-factorization method. $3x_1 - x_2 = -1, \;\;\; -x_1+3x_2-x_3 = 7, \;\;\; -x_2+3x_3 =7$. [4]

\question Determine whether the matrix is ill-conditioned or not.$\small A=   \begin{bmatrix}
    25  & 24 \\
    66 & 78 \\
  \end{bmatrix}$. [2]

\question Using Lagrange's method estimate the value of $y(202)$ from the following data: \\[1mm]
  $(200, 5.2983), (203,5.3132), (206,5.3278), (210,5.3471)$. [4]

\end{questions}
\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: