### 2025-06-12 question_papers/aima203_first_in_sem.tex
question paper

\documentclass[12pt]{exam}
\usepackage{preamble_question_small}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
 {\bfseries  { Kathmandu University}} \\[-1mm]
 First In-Semester Exam-2025\\[-2mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\begin{minipage}{0.70\linewidth}
  \small
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 1mm
\rule{\textwidth}{1pt}
\vskip -5mm
\begin{questions}

\question Round off the digits 45.321 and 12.987 upto two decimal places and find the error in their product after rounding off. [2]

\question Write the rate of convergence of Newton Raphson's method. Find a root of \(f(x)=xe^x-1=0\), correct to \(2\) decimal places. [1+3]

\question Find a real root of the equations $y^2-5y+4=0$ and $3yx^2-10x+7=0$ correct to 3 decimal places, using the method of iteration. [4]

\end{questions}

\vspace{4cm}
\begin{center}
 {\bfseries  {Kathmandu University}} \\[-1mm]
 First In-Semester Exam-2025\\[-2mm]
 {\small \textbf{Department of Artificial Intelligence, Panchkhal}}
\end{center}
\begin{minipage}{0.70\linewidth}
  \small
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 1mm
\rule{\textwidth}{1pt}

\begin{questions}

\question Round off the digits 45.321 and 12.987 upto two decimal places and find the error in their product after rounding off. [2]

\question Write the rate of convergence of Newton Raphson's method. Find a root of \(f(x)=xe^x-1=0\), correct to \(2\) decimal places. [1+3]

\question Find a real root of the equations $y^2-5y+4=0$ and $3yx^2-10x+7=0$ correct to 3 decimal places, using the method of iteration. [4]

\end{questions}

  \clearpage

  \vspace{4cm}

  \begin{questions}
      \setcounter{question}{3}
\question Solve using LU-factorization method. $3x_1 - x_2 = -1, \;\;\; -x_1+3x_2-x_3 = 7, \;\;\; -x_2+3x_3 =7$. [4]

\question Determine whether the matrix is ill-conditioned or not.$\small A=   \begin{bmatrix}
    5  & 4 \\
    6 & 8 \\
  \end{bmatrix}$. [2]

\question Using Lagrange's method estimate the value of $y(202)$ from the following data: \\[2mm]
  $(200, 5.2983), (203,5.3132), (206,5.3278), (210,5.3471)$. [4]

 \begin{center}
    \bfseries OR
  \end{center}
  Using Newton's Interpolation formula estimate : $y(10)$ from \\
  $(1,24), \;\;\; (3,120),\;\;\; (5,366), \;\;\; (7,720)$

  \vspace{9cm}
  \setcounter{question}{3}
 \question Solve using LU-factorization method. $3x_1 - x_2 = -1, \;\;\; -x_1+3x_2-x_3 = 7, \;\;\; -x_2+3x_3 =7$. [4]

\question Determine whether the matrix is ill-conditioned or not.$\small A=   \begin{bmatrix}
    5  & 4 \\
    6 & 8 \\
  \end{bmatrix}$. [2]

\question Using Lagrange's method estimate the value of $y(202)$ from the following data: \\[1mm]
  $(200, 5.2983), (203,5.3132), (206,5.3278), (210,5.3471)$. [4]
  \begin{center}
    \bfseries OR
  \end{center}
  Using Newton's Interpolation formula estimate : $y(10)$ from \\
  $(1,24), \;\;\; (3,120),\;\;\; (5,366), \;\;\; (7,720)$

\end{questions}
\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: