### 2025-06-12 question_papers/aima203_mcq_first_in_sem.tex

\documentclass[12pt]{exam}
\usepackage{preamble_question}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
  {\bfseries  {\large Kathmandu University}} \\
  First In-semester Examination-2025 \\[-2mm]
    \textbf{\small Department of Artificial Intelligence, Panchkhal}
\end{center}
\vspace{2mm}

\begin{minipage}{0.70\linewidth}
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: II \\
    Time: 15 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 10
  \end{flushleft}
\end{minipage}
\vspace{-8mm}
\begin{center}
  \rule{\textwidth}{1pt}
  Name: \hspace{5cm}  Roll No: \hspace{3cm} Marks-Scored:
  \vskip -3mm
\rule{\textwidth}{1pt}
\end{center}

\begin{center}
  SECTION''A'' \hspace{5mm} [10 Q \(\times\) 1 = 10 marks] (\textit{Do any ten}).
\end{center}
Fill in the blanks by writing the most appropriate word(s) or symbol(s).
\begin{questions}

\question Write the number of significant digits in 20300.00  \rule{5cm}{0.15mm}.
  \vspace{2mm}

  \question Write the precision and accuracy of 45.2030  \rule{5cm}{0.15mm}.

  \question Find the upper limit error in rounding a number to $4$ decimal places.\rule{4cm}{0.15mm}.

    \question What is the maximum number of positive real roots of $x^5-x^4+3x^3-x^2-x+2$? \rule{7cm}{0.15mm}.

    \question Which error: Truncation error or rounding off error in Newton Raphson's method? \rule{5cm}{0.15mm}.

    \question Which one converges faster bisection method or secant method? \rule{5cm}{0.15mm}.

    \question Write the disadvantage of ``Lagrange's Interpolation method'': \rule{5cm}{0.15mm} \\[1mm]
      \rule{10cm}{0.15mm}.

    \question Write the advantage of ``Secant method over Newton Rapshon's method'': \\[1mm]
      \rule{12cm}{0.15mm}.

    \question Write the formula for $\Delta y_1$ and $\nabla y_1$: \rule{7cm}{0.15mm}.

    \question What is the difference of interpolation and extrapolation. \rule{5cm}{0.15mm} \\[1mm]
      \rule{12cm}{0.15mm}.

 \question Calculate the number of steps taken by Bisection method to get the accuracy less than $10^{-3}$.\rule{8cm}{0.15mm}.
\end{questions}

\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: