### 2025-06-16 question_papers/aima104_mcq_first_in_sem.tex

\documentclass[12pt]{exam}
\usepackage{preamble_question}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
  {\bfseries  {\large Kathmandu University}} \\
  First In-semester Examination-2025 \\[-2mm]
    \textbf{\small Department of Artificial Intelligence, Panchkhal}
\end{center}
\vspace{2mm}

\begin{minipage}{0.70\linewidth}
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: I \\
    Time: 15 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 10
  \end{flushleft}
\end{minipage}
\vspace{-8mm}
\begin{center}
  \rule{\textwidth}{1pt}
  Name: \hspace{5cm}  Roll No: \hspace{3cm} Marks-Scored:
  \vskip -3mm
\rule{\textwidth}{1pt}
\end{center}

\begin{center}
  SECTION''A'' \hspace{5mm} [10 Q \(\times\) 1 = 10 marks] (\textbf{\textit{Do any ten}}).
\end{center}
Fill in the blanks by writing the most appropriate word(s) or symbol(s).
\begin{questions}
\question Give an example of a vector and its scalar. \rule{7cm}{0.15mm}.
\question Give an example of a subspace of $\mathbb{R}^2$ : \rule{8cm}{0.15mm}.

\question Give an example of $3 \times 4 $ row echelon for matrix. \rule{7cm}{0.15mm}.

\question Find the rank of the matrix $\displaystyle \begin{bmatrix}
    1 & 3 & 5 \\
    2 & 6 & 10 \\
    4  & 12 & 20
\end{bmatrix}.$ \rule{7cm}{0.15mm}.

\question Determine whether the system has unique solution, no unique solution or no solution. $ \left[ \begin{matrix}
    1 & 0  \\
    0 & 1  \\
    0 & 0
\end{matrix} \;\; \vline \;\;
\begin{matrix}
    0\\ 0\\ -1
\end{matrix} \right]$ \rule{7cm}{0.15mm}.

\question Write a basis of the vector space $\{0\}$. \rule{7cm}{0.15mm}.

\question Suppose a $4 \times 7$ matrix $A$ has four pivot columns. Is $Col(A) = \mathbb{R}^4$? Is $Nul(A) = \mathbb{R}^3$ ? \rule{12cm}{0.15mm}.

\question Is the set of vectors linearly dependent?$\{(1,2, 4) \;\; (2,4,8)\}$ \rule{4cm}{0.15mm}.

\question What are the eigenvalues of a triangular matrix? \rule{7cm}{0.15mm}.

\question Explain why there are infinitely many eigenvalues associated with an eigenvalue.\\[1mm]
  \rule{12cm}{0.15mm}.

\question If $Ax=2x$ then, $AAx=?$ \rule{7cm}{0.15mm}.

\question Can you explain the case where zero is an eigenvalue of a matrix? \\[1mm] \rule{12cm}{0.15mm}.
\end{questions}

\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: