### 2025-06-28 lecturenotes_aima203/ch6_aima203.tex
ch 6

\documentclass[aima203_lecturenotes_ku.tex]{subfiles}

\setcounter{chapter}{5}
\begin{document}
\chapter{Numerical Differentiation and Numerical Integration}

In this chapter we shall be concerned with the problems of numerical differentiation and integration. We shall derive the formula to compute the following when only tabulated values of the function are known but the explicitly nature of the function is not known. Such scenario occurs in engineering in case of experimental data:

\begin{itemize}
\item $\displaystyle \frac{dy}{dx}, \; \frac{d^2y}{dx^2},\;...$ for any value of $x$ in $[x_0, x_n]$, and

\item $\displaystyle \int_{x_0}^{x_n} \; y\,dx$
\end{itemize}

\section{Numerical Differentiation}
The general method for deriving the numerical differentiation formulae is to differentiate the interpolating polynomial. Hence, corresponding to each of the Interpolating formula derived, we may derive a formula for the derivative.

\subsection{Newton's forward difference formulae}
The Newton's forward difference formula is:
\begin{equation}
  \label{newtforward6}
 \begin{gathered}
  y_n(x) = y_0 + u \Delta y_0 + \frac{u(u-1)}{2!}\, \Delta ^2 y_0 + \frac{u(u-1)(u-2)}{3!}\, \Delta ^3 y_0 + ... \\[1mm]
  + \frac{u(u-1)(u-2)...(u-(n-1))}{n!}\, \Delta ^n y_0
\end{gathered}
\end{equation}
where, $x=x_0 + uh$ and $h = x_{i+1} - x_{i}$. Differentiating ~\ref{newtforward6} with respect to $x$,

\begin{equation}
  \label{newtforwarddif}
  \begin{aligned}
    \frac{dy}{dx} =\frac{dy}{du} \frac{du}{dx} = \frac{1}{h} \left [ \Delta y_0 + \frac{2u -1}{2}\, \Delta^2\,  y_0 + \frac{3u^2-6u + 2}{6}\, \Delta^3 \, y_0 + ... \right ]
  \end{aligned}
\end{equation}
Differentiating ~\ref{newtforwarddif} with respect to $x$, we get,
\begin{equation}
  \label{newtforwarddif2}
  \begin{aligned}
    \frac{d^2y}{dx^2} = \frac{1}{h^2} \left [ \Delta^2 y_0 + \frac{6u -6}{6}\, \Delta^3 \,  y_0 + \frac{12u^2-36u + 22}{24}\, \Delta^4 \, y_0 + ... \right ]
  \end{aligned}
\end{equation}
These formulae are used for \textit{non-tabular} values of $x$. For tabular values of $x$, the formulae take simpler form. For $x=x_0$, we have $u=0$, and using this we can find the relations of $\displaystyle \left [ \frac{dy}{dx} \right ]  _{x=x_0} and \left [\frac{d^2y}{dx^2} \right ] _{x=x_0}$ which is in a simpler form.
\newpage
\begin{equation}
  \label{newtforw1x}
  \left [ \frac{dy}{dx} \right ] _{x=x_n} = \frac{1}{h} \left [\Delta \, y_n - \frac{1}{2} \Delta ^2 \, y_n + \frac{1}{3} \Delta ^3 \, y_n - ... \right ]
\end{equation}
and
\begin{equation}
  \label{newtforw2x}
  \left [ \frac{d^2y}{dx^2} \right ] _{x=x_n} = \frac{1}{h^2} \left [\Delta^2 \, y_n - \Delta ^3 \, y_n + \frac{11}{12} \Delta ^4 \, y_n - \frac{5}{6} \Delta ^5 \, y_n + ... \right ]
\end{equation}

\subsection{Newton's backward difference formula}
In a similar way, different formulae can be derived by starting with other interpolation formulae. Thus, Newton's backward difference formula gives
\begin{equation}
  \label{newtbackwarddif}
  \left [ \frac{dy}{dx} \right ] _{x=x_n} = \frac{1}{h} \left [\nabla \, y_n + \frac{1}{2} \nabla ^2 \, y_n + \frac{1}{3} \nabla ^3 \, y_n + ... \right ]
\end{equation}
and
\begin{equation}
  \label{newtbackwarddif2}
  \left [ \frac{d^2y}{dx^2} \right ] _{x=x_n} = \frac{1}{h^2} \left [\nabla^2 \, y_n + \nabla ^3 \, y_n + \frac{11}{12} \nabla ^4 \, y_n + \frac{5}{6} \nabla ^5 \, y_n + ... \right ]
\end{equation}
\begin{remark}
   If the $x$-values of the data points are equally spaced then it is common to use Newton's interpolation.
\end{remark}
\subsection{Exercise}
Find the $dy/dx$ and $d^2y/dx^2$ at $x=1.2, \; x=1.6, \; x=2.2$ from the data: \\[1mm]
$(1, 2.7183), \; (1.2, 3.3201), \; (1.4, 4.0522), \; (1.6, 4.9530), \; (1.8, 6.0496), \; (2.0, 7.3891), \; (2.2, 9.0250)$.
\section{Numerical Integration}
The general problem of numerical integration may be stated as: Given a set of data points of a function $y=f(x)$, where $f(x)$ is not known explicitly, it is required compute the value of the definite integral $$I = \int_a^b \; y\,dx$$ As with the case of numerical differentiation, one replaces $f(x)$ by an interpolating polynomial $\phi(x)$ and obtains, on integration, an approximate value of the definite integral. Thus, different integration formulae can be obtained depending upon the type of the interpolation formula used. Approximating $y$ by Newton's forward difference formula, we obtained
$$ I = \int_{x_0}^{x_n} \; \left [ y_0 + p \Delta y_0 + \frac{p(p-1)}{2}\, \Delta^2 y_0 + \frac{p(p-1)(p-2)}{6}\, \Delta ^3 y_0 + ... \right ] \, dx$$
Since $x=x_0 +ph$, $dx = hdp$. When $x=x_0$, $p=0$ and $x=x_n$, $p=n$. Then,
$$ I = \int_{0}^{n} \; \left [ y_0 + p \Delta y_0 + \frac{p(p-1)}{2}\, \Delta^2 y_0 + \frac{p(p-1)(p-2)}{6}\, \Delta ^3 y_0 + ... \right ] \, h\,dp$$
Integrating we get,
\begin{equation}
  \label{numint}
  \begin{aligned}[b]
    I &= h \left [ py_0 + \frac{p^2}{2}\, \Delta y_0 + \left ( \frac{p^3}{3} - \frac{p^2}{2} \right ) \, \frac{\Delta^2 y_0}{2} +  \left ( \frac{p^4}{4} - p^3 + p^2 \right ) \frac{\Delta ^3 \, y_0}{6} + ... \right ]_0^n \\[1mm]
    &= nh \left [ y_0 + \frac{n}{2}\, \Delta y_0 + \left ( \frac{2n^2-3n}{12} \right ) \, \Delta^2\, y_0 +  \left ( \frac{n^3 - 4n^2 +4n}{24} \right ) \, \Delta ^3 \, y_0 + ... \right ]
  \end{aligned}
\end{equation}

This relation ~\ref{numint} is considered to be a general formula in the variable $n$. For a particular value of $n$ we get a particular formula. For example for $n=1$ we get the famous Trapezoidal rule and for $n=2$ and $n=3$ we get Simpson's 1/3 rule and Simpson's 3/8 rule respectively.

\subsection{Trapezoidal Rule}
Setting $n=1$ in the general formula ~\ref{numint}, all differences higher than the first order will become zero and we obtained \\[1mm]
$\displaystyle \int_{x_0}^{x_1} \; y\,dx = h \left [ y_0 + \frac{1}{2} \Delta y_0 \right ] = h \left [ y_0 + \frac{1}{2}(y_1 -y_0) \right ] = \frac{h}{2}\, (y_0 + y_1).$ Similarly we can obtain \\[5mm] $\int_{x_1}^{x_2} \; y\,dx, ..... \int_{x_{n-1}}^n \; y\,dx $. Then we have,
\begin{equation}
  \label{trap}
  \begin{aligned}[b]
    \int_{x_0}^{x_n}\; y\,dx &= \int_{x_0}^{x_1} \; y\,dx + \int_{x_1}^{x_2} \; y\,dx + .... + \int_{x_{n-1}}^{x_n} \; y\,dx \\[1mm]
                             &= \frac{h}{2} \, (y_0 + y_1) + \frac{h}{2} \, (y_1 + y_2) + .... + \frac{h}{2} \, (y_{n-1} + y_n) \\[1mm]
                             &= \frac{h}{2} \, \{ y_0 + 2(y_1 + y_2 + ... + y_{n-1}) + y_n \}
  \end{aligned}
\end{equation}
\textbf{Question:} Illustrate the trapezoidal method geometrically.

\subsection{Simpson's 1/3-Rule}
Setting $n=2$ in the general formula ~\ref{numint}, all differences higher than the second will become zero and we obtained \\[1mm]
$\displaystyle \int_{x_0}^{x_2} \; y\,dx = 2h \left [ y_0 + \Delta \, y_0 + \frac{1}{6} \Delta^2 \, y_0 \right ] = \frac{h}{3}\, (y_0 + 4y_1 + y_2).$ Similarly we can obtain \\[5mm] $\int_{x_2}^{x_4} \; y\,dx, ..... \int_{x_{n-2}}^n \; y\,dx $. Then we have,
\begin{equation}
  \label{simp1}
  \begin{aligned}[b]
    \int_{x_0}^{x_n}\; y\,dx &= \int_{x_0}^{x_2} \; y\,dx + \int_{x_2}^{x_4} \; y\,dx + .... + \int_{x_{n-2}}^{x_n} \; y\,dx \\[1mm]
                             &= \frac{h}{3} \, (y_0 +4y_1 + y_2) + \frac{h}{3} \, (y_2 + 4y_3 + y_4) + .... + \frac{h}{3} \, (y_{n-2} + 4y_{n-1} + y_n) \\[1mm]
                             &= \frac{h}{3} \, \{ y_0 + 4(y_1 + y_3 + ... + y_{n-1}) + 2 (y_2 + y_4  ... + y_{n-2} ) + y_n \}
  \end{aligned}
\end{equation}

\begin{remark}
  It should be noted that this rule requires the division of the whole range into an even number of sub-intervals of width $h$.
\end{remark}

\subsection{Simpson's 3/8-Rule}
Setting $n=3$ in the general formula ~\ref{numint}, all differences higher than the third will become zero and we obtained \\[1mm]
$\displaystyle \int_{x_0}^{x_3} \; y\,dx = 3h \left [ y_0 + \frac{3}{2} \Delta \, y_0 + \frac{3}{4} \Delta^2 \, y_0 + \frac{1}{8} \Delta ^3 \, y_0 \right ] = \frac{3h}{8}\, (y_0 + 3y_1 + 3y_2+y_3).$ \\ Similarly we can obtain \\[5mm] $\int_{x_3}^{x_6} \; y\,dx, ..... \int_{x_{n-3}}^n \; y\,dx $. Then we have,
\begin{equation}
  \label{simp2}
  \begin{aligned}[b]
    \int_{x_0}^{x_n}\; y\,dx &= \int_{x_0}^{x_3} \; y\,dx + \int_{x_3}^{x_6} \; y\,dx + .... + \int_{x_{n-3}}^{x_n} \; y\,dx \\[1mm]
                             &= \frac{3h}{8} \, (y_0 +3y_1 + 3y_2 + y_3) + \frac{3h}{8} \, (y_3 + 3y_4 + 3y_5 + y_6) + .... + \frac{3h}{8} \, (y_{n-3} + 3y_{n-2} + 3y_{n-1} + y_n) \\[1mm]
                             &= \frac{3h}{8} \, \{ y_0 + 3(y_1 + y_2 + y_4 + y_5 + ... + y_{n-1}) + 2 (y_3 + y_6  ... + y_{n-3} ) + y_n \}
  \end{aligned}
\end{equation}

\subsection{Exercise}
\begin{enumerate}
\item For the given $(x,y)$ data points: Evaluate $\int_0^2 \, ydx$ using the trapezoidal rule.\\  $(0.0,0.399),\, (0.5,0.352),\,(1.0,0.242),\, (1.5,0.129),\,(2.0,0.054)$.

\item Evaluate using Simpson's $1/3$ rule:
  \begin{enumerate}
  \item $\displaystyle \int_1^3 \, \frac{1}{x}\,dx$  with $8$ stripes.
  \item $\displaystyle \int_0^{\pi /2} \, \sqrt{sinx}\, dx$ with $h= \pi /12$.

  \item $\displaystyle \int_3^7 \, x^2logx\, dx $ with $h= 1$.
    \item $\displaystyle \int_0^1 \, e^{-x^2} \, dx$ with $h=0.1$
  \end{enumerate}

\item Evaluate using Simpson's $3/8$ rule: $\displaystyle \int_0^1 \, \frac{dx}{1+x}$ with $h=1/6$.

\item Compute the values of $I = \displaystyle \int_0^1 \, \frac{dx}{1+x^2}$ using the trapezoidal rule with $h=0.5, \; 0.25, \; 0.125$. Then obtain a better estimate using \textbf{Romberg's method}.

\item Compute $\displaystyle I\left (\frac{\pi}{4} \right ), \; I\left (\frac{\pi}{8} \right ), \; I\left (\frac{\pi}{8}, \frac{\pi}{4} \right )$ for : $I = \displaystyle \int_0^{\pi/4} \, cos^2x\, dx$.

\item Using Trapezoidal rule and Simpson's $1/3$ rule evaluate the double integral $$\int_{_2}^2 \, \int_0^4 \; (x^2 -xy + y^2) \; dx\, dy.$$

\end{enumerate}
\subsection{Boole's and Weddle's Rules, \textit{Extra material}}
To incorporate the fourth order difference as well we need $5$ data points $x_0, ..., x_4$ and we need to integrate from $x_0$ to $x_4$. Then we obtain Boole's formula:
\begin{equation}
  \label{Boole's}
  \displaystyle \int_{x_0}^{x_4} \; y\,dx = \frac{2h}{45}\, (7y_0 + 32y_1 + 12y_2+32y_3+7y_4)
\end{equation}
If we incorporate differences u to six order then we obtain Weddle's formula as follows:
\begin{equation}
  \label{Weddle's}
  \displaystyle \int_{x_0}^{x_6} \; y\,dx = \frac{3h}{10}\, (y_0 + 5y_1 + y_2+6y_3+y_4+5y_5+y_6)
\end{equation}

\subsection{Romberg Integration}
This method improves the approximation obtained from the other methods. Here we describe the Romberge method to improve the results obtained from the Trapezoidal method. \\[2mm]
Suppose the integral $\displaystyle \int_a^b \, y\,dx$ is evaluated by the Trapezoidal rule with two different subintervals of widths $h_1$ and $h_2$, to obtain the approximate values $I_1$ and $I_2$, respectively. Romberg's method uses this two approximations: $I_1$ and $I_2$ to obtain the next better approximations which is given by
\begin{equation}
  \label{romberg}
I_3 = \frac{I_1h_2^2-I_2h_1^2}{h_2^2 - h_1^2}
\end{equation}

\subsection{Gaussian Quadrature}
The methods in the previous sections are based upon the values of function a equally-spaced points of the interval. Gauss derived a formula for which the points in the interval \textbf{need not be equally spaced}.
The Gaussian formula is expressed as:
\begin{equation}
  \label{gaussian}
\int_{-1}^{1} \, F(u)\, du = \sum_{i=1}^n \; W_i\, F(u_i)
\end{equation}
where $W_i$ and $u_i$ are called the \textit{weights} and \textit{abscissae}, respectively. It can be shown that the $u_i$ are the zeros of the $(n+1)^{th}$ \textbf{Legendre} polynomial $P_{n+1}(u)$ which can be generated usingthe recurrence relation:
\begin{equation*}
  (n+1)\,P_{n+1}(u) = (2n+1)\,u\,P_{n}(u) -n\,P_{n-1}(u)
\end{equation*}
where $P_0(u)=1$ and $P_1(u)=u$. It can also be shown that corresponding weights $W_i$ are given by
\begin{equation*}
  W_i =\int_{-1}^{1}\; \prod_{j=0, j\neq i}^n \; \frac{u-u_j}{u_i - u_j}\, du
\end{equation*}
However the weights and the abscissae are extensively tabulated for different values of $n$. We list the weights and the abscissae for values of $n$ up to $n=6$. \\[1mm]
\textbf{Change of Limits}
In the general case, the limits of the integral have to be changed to that of ~\ref{gaussian} by means of the following variable transformation:
\begin{equation*}
  x = \frac{1}{2} \, u(b-a) + \frac{1}{2}\, (a+b)
\end{equation*}

\begin{figure}[h]
  \centering
\includegraphics[width=12cm, height=9cm]{gaussian.png}
\end{figure}
\subsection{Exercise}
\begin{enumerate}
\item Derive the Gauss quadrature formula for $n=2$ and apply it to evaluate the integral $\displaystyle \int_{-1}^1 \, \frac{dx}{1+x^2}$.

\item Use the three point Gauss formula to evaluate the integral $\displaystyle \int_0^1 \, \frac{dx}{1+x}$.
\end{enumerate}
\section{Double Integration}
Formulae for the evaluation of double integral can be obtained by repeatedly applying the trapezoidal and Simpson's rules derived earlier.
\subsection{Trapezoidal Rule}
\begin{equation*}
  I = \int_{y_j}^{y_{j+1}}\; \int_{x_i}^{x_{i+1}} \; f(x,y)\;dx\,dy
\end{equation*}
where, $x_{i+1} = x_i + h$ and $y_{j+1} = y_j + k$.
By repeated application of trapezoidal rule to we get,
\begin{equation}
  \label{trapdouble}
  \begin{aligned}
    I &= \frac{h}{2} \,\int_{y_j}^{y_{j+1}}\; [f(x_i,y) + f(x_{i+1},y) ]\;dx\,dy \\[1mm]
      &= \frac{hk}{4} \, [f(x_i,y_j) + f(x_{i+1},y_j) + f(x_i,y_{j+1}) + f(x_{i+1},y_{j+1}) ]\;dx\,dy \\[1mm]
      &= \frac{hk}{4} \, [f_{i,j} + f_{i+1, j} + f_{i,j+1} + f_{i+1,j+1} ]\;dx\,dy
  \end{aligned}
\end{equation}

\subsection{Simpson's Rule}
\begin{equation*}
  I = \int_{y_{j-1}}^{y_{j+1}}\; \int_{x_{i-1}}^{x_{i+1}} \; f(x,y)\;dx\,dy
\end{equation*}
where, $x_{i+1} = x_i + h$ and $y_{j+1} = y_j + k$.
By repeated application of Simpson's rule to we get,
\begin{equation}
  \label{simpdouble}
  \begin{aligned}
    I &= \frac{h}{3} \,\int_{y_{j-1}}^{y_{j+1}}\; [f(x_{i-1},y) + 4f(x_i,y) + f(x_{i+1},y) ]\;dx\,dy \\[1mm]
      &= \frac{hk}{9} \, [f(x_{i-1},y_{j-1}) + 4f(x_{i-1},y_j) + f(x_{i-1},y_{j+1}) \\[1mm]
    & 4\{f(x_i,y_{j-1}) + 4f(x_i,y_j) + f(x_i,y_{j+1})\} + f(x_{i+1},y_{j-1}) + 4f(x_{i+1},y_j) + f(x_{i+1},y_{j+1})]
  \end{aligned}
\end{equation}
\vspace{1cm}

\begin{figure}[h]
  \centering
\includegraphics[width=15cm, height=11cm]{doubleint.png}
\end{figure}

\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: