### 2025-08-26 lecturenotes_aima203/title_aima203.tex

\vspace*{2cm}
{\begin{center}
  {\Huge
    {\bfseries {\thetitle}}} \\
        \vspace{5mm}
        {\large AIMA 203}
  \vspace{3cm}

\begin{figure}[h]
	\centering
	\includegraphics[height=4cm,width=4cm]{images/Kulogo.png}
\end{figure}

\vspace{1cm}

	{\large \bfseries Department of Artificial Intelligence, \\[1mm]
 Kathmandu University, Dhulikhel, Kavre}

\vspace{2cm}
        \textit{Lecture notes}
	by\\
	{\bfseries  \theauthor}

\vspace{1cm}

    \textbf{Text Book}: S.S. Sastary, Introductory Methods of Numerical Analysis, Prentice Hall of India.
    \vspace{1.5cm}

{\thedate}

\thispagestyle{empty}
\end{center}}

\clearpage