### 2025-08-26 question_papers/aima104_mcq_second_in_sem.tex

\documentclass[12pt]{exam}
\usepackage{preamble_question}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}

\begin{center}
  {\bfseries  {\large Kathmandu University}} \\
  Second In-semester Examination-2025 \\[-2mm]
    \textbf{\small Department of Artificial Intelligence}
\end{center}
\vspace{2mm}

\begin{minipage}{0.70\linewidth}
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: I \\
    Time: 15 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 203 \\
    Semester: II \\
    F.M. : 10
  \end{flushleft}
\end{minipage}
\vspace{-8mm}
\begin{center}
  \rule{\textwidth}{1pt}
  Name: \hspace{5cm}  Roll No: \hspace{3cm} Marks-Scored:
  \vskip -3mm
\rule{\textwidth}{1pt}
\end{center}

\begin{center}
  SECTION''A'' \hspace{5mm} [10 Q \(\times\) 1 = 10 marks] (\textbf{\textit{Do any ten}}).
\end{center}
Fill in the blanks by writing the most appropriate word(s) or symbol(s).
\begin{questions}
\question Is $
  \begin{bmatrix}
    1 \\ 4
  \end{bmatrix}$ an eigenvector of $\begin{bmatrix} -3 & 1  \\  -3 & -8  \\ \end{bmatrix}$ ? \rule{7cm}{0.15mm}.
\question Give an example of a subspace of $\mathbb{R}^2$ : \rule{8cm}{0.15mm}.

\question What is the sufficient condition for diagonalization of a matrix? \rule{4cm}{0.15mm} \\[1mm]
\rule{9cm}{0.15mm}.

\question Write the relationship between algebraic multiplicity and geometric multiplicity of an eigenvalue. \rule{8cm}{0.15mm}.

\question Compute the distance between: $u=
  \begin{bmatrix}
    -1 \\ 2
  \end{bmatrix}, \vspace{4mm}
   v=
   \begin{bmatrix}
     4 \\6
   \end{bmatrix}$ \rule{7cm}{0.15mm}.

\question Find the angle between $u$ and $v$ given above. \rule{7cm}{0.15mm}.

\question Let $W$ be a subspace of $V$. What is the intersection between $W$ and $W^{\perp}$? \rule{3cm}{0.15mm}.

\question In QR factorization, what is the nature of $R$? \rule{7cm}{0.15mm}.
\vspace{2mm}

\question Give an example of Hermitian matrix. \rule{7cm}{0.15mm}.

\question Find orthogonal projection of $u$ onto $v$ for the given $u$ and $v$ in question 5. \rule{3cm}{0.15mm}. \\[1mm]
\rule{7cm}{0.15mm}.
\end{questions}

\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: