### 2025-08-26 question_papers/aima104_second_in_sem.tex

\documentclass[12pt]{exam}
\usepackage{preamble_question_small}
\usepackage[normalem]{ulem}
\renewcommand\ULthickness{2pt}
\setlength\ULdepth{1.3ex}

\begin{document}
\begin{center}
 {\bfseries  {\large Kathmandu University}} \\
 Second In-Semester Exam-2025\\[-1mm]
 {\small \textbf{Department of Artificial Intelligence}}
\end{center}

\begin{minipage}{0.70\linewidth}
  \begin{flushleft}
    Level: B.Tech Artificial Intelligence \\
    Year: I \\
    Time: 60 minutes
  \end{flushleft}
\end{minipage} \hfill
\begin{minipage}{0.25\linewidth}
  \begin{flushleft}
    Course: AIMA 104 \\
    Semester: II \\
    F.M. : 20
  \end{flushleft}
\end{minipage}
\vskip 2mm
\rule{\textwidth}{1pt}
\vskip 2mm
\begin{questions}
\question Find the all the eigen-spaces of $A= \begin{bmatrix}
      4 & 1 & 1 \\
      2 & 4 & 1 \\
      0 & 1 & 4
\end{bmatrix}$? [4]

\question Diagonalize the matrix, if possible: $
   \begin{bmatrix}
     4 & 2 & 2 \\
     2 & 4 & 2 \\
     2 & 2 &4
   \end{bmatrix}
   $. [2]

\question Give an example of similar matrices [2].

\question For a subset $W$ of a vector space $V$, show that $W^{\perp}$ is a subspace $V$. [2]

\question Let $\{u_1, ..., u_p\}$ be an orthogonal basis for a subspace $W$ of $\mathbb{R}^n$. Prove that, for each $y$ in $W$, the coordinates of $y$ with respect to the orthogonal basis : $y = c_1u_1 + \; ... \; c_n u_n$ are given by $\displaystyle c_j = \frac{y.u_j}{u_j.u_j}$. [4]

\question Find an orthogonal basis for the column space of the matrix $\begin{bmatrix}
      -1 & 6 & 6 \\
      3 & -8 & 3 \\
      1 & -2 & 6 \\
      1 & -4 & -3
    \end{bmatrix}$, given that the columns are linearly independent. [4]
\end{questions}
\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End: